Spin canonical divisors of log stacky curves

### 2015-06-27 stacky_write_up/spin_cring.tex
improved intro, added some comments

Following Voight and Zureick-Brown \cite{vzb:stacky}, we generalize Petri's theorem in many
directions. A divisor $\Delta \in \di X$ is a {\bf log divisor} if
$\Delta = \sum_{i=1}^{n}P_i \in \di X,$ with $P_i$ distinct points of $X$. Let
$\sx$ denote a tame separably rooted curve over a
field $k$ so that $X$ is the coarse space of $\sx$, as defined in
\cite[Section 5.2]{vzb:stacky}, with $\di \sx = \di X \oplus \left(\bigoplus_{i
= 1}^r \langle \frac{1}{e_i}P_i \rangle \right),$ so that $e_i \in \BZ$. Let $K_
\sx = K_X + \sum_{i = 1}^{r}\frac{e_i - 1}{e_i}P_i.$ In this case, we
say $\sx$ has {\bf stabilizer order} $e_i$ at $P_i$. Call $\halfcan
\in \di \sx$ a {\bf log half-canonical divisor}, also known as a theta-characteristic, on the {\bf log stacky curve} $(\sx,\Delta)$ if
$2\halfcan \sim K_\sx + \Delta$. If $X$ has genus $g$ and $\deg \Delta =\delta,$ then we
say $\sx$ has \textbf{signature} $\sigma = (g; e_1, \ldots, e_r; \delta)$. A
{\bf log spin curve} is a triple $(\sx, \Delta, \halfcan)$ where $(\sx,\Delta)$
is a log stacky curve and $\halfcan$ is a log half-canonical divisor on $(\sx,\Delta)$.
The central object of study in this paper is the {\bf log spin
canonical ring} of $(X,\Delta,L)$ is
\begin{align*}
	R(\sx,\Delta,L) = \bigoplus_{d \geq 0} H^0(\sx,\sco(L)^{\otimes d})
\end{align*}

@@

\todo{Somewhere in the paper, we should probably give an example showing the generators and relations are sharply bounded at degree e, 2e. Robin: Aren't the base cases suitable examples?}

\section{Background}
\label{sec:background}
Look at Spencer-Kodaira \cite{kodaira:complex-manifolds}.
Include the following:
\begin{enumerate}
	\item Fix notation $\delta$ for $\deg \Delta$, $g$ for genus.
	\item While the remainder paper could be writen using stack-theoretic terms, as is done in \cite{vzb:stacky}, we shall adopt the simpler language of $\BQ$ divisors, thinking of $\sx$ as the underlying scheme $X$ with fractional points.
	\item Include explicit definitions for what we are assuming about a curve, i.e. smooth, projective, etc. Also, possibly discuss the relations to stacks, either here in or in the introduction. Make sure to make clear the assumption that we will work over an algebraically closed field through tensoring up.
	\item Notation for Coarse space
	\item Notation for Stacky points with associated stabilizer orders (I.e. as those points for which the divisor can have rational coefficients with denominator equal to the stabilizer order)
	\item Note that this result is independence of base field
	\item background on Gr\"{o}bner bases
	\item Notation $h^0 = \dim_k H^0$.
	\item Why we have these fractions $\frac{e_i - 1}{2e_i}$ and why $e_i$ must be odd.
	\item Include notation for $X$ as coarse space and $L_X$ the corresponding log half canonical divisor.
	\item Notation for Gr\"{o}bner bases and orderings (include grlex, grevlex, and block order, as defined in \cite[Definition 8.1.1]{vzb:stacky})
	\item Make sure to standardize the notation $\sx,\sx'$ for stacky curves, $L,L'$ for half canonical divisors, and $X,X'$ for (non-stacky) curves.
\end{enumerate}

@@

In Lemma \ref{lem:sat-1} we provide an inductive method to strengthen
O'Dorney's result \todo{cite theorem 8} on $\BP^1_k$ to divisors $D = \sum_{i=1}^n \frac{\alpha_i
}{\beta_i} P_i$ for points $P_1, \ldots, P_n$ on a curve $X$ such that $H^
0(X, \lfloor D \rfloor) > 0$; our result implies that $R_D$ is generated in
degrees at most $e = \max(\beta_i)$ with relations generated in
degrees at most $2e$.  Furthermore, given additional constraints on $H^0(
X, \lfloor dD \rfloor)$ for various values of $d$, we can apply the
same Riemann-Roch arguments in arbitrary genus. 
To do this, we extend an inductive result by Voight and
Zureick-Brown on stacky canonical rings to arbitrary $\BQ$-
divisors. \todo{cite theorem 8.3.1}

@@

As suggested by the previous lemma, we will need an ordering on monomials, and the following definition will prove useful in the ensuing lemma.
\begin{defn}
We equip the graded ring $k[x_1, \ldots x_m]$, with weight $a_i$ and pole order $-\ord_P(x_i)$ attached to $x_i$, the graded $P$ reverse lexicographic order {\bf{g-P-revlex}} if for $f=\prod_{i=1}^m {x_i}^{q_i}\ne g=\prod_{i=1}^m {x_i}^{r_i}$, $f\prec g$ iff one of the following holds
\begin{enumerate}
\item $\deg(f) < \deg(g)$
\item $\deg(f) = \deg(g)$ and $-\ord(f) > -\ord(g)$
\item $\deg(f) = \deg(g)$, $-\ord(f) = -\ord(g)$, and for $i$ maximal such that $q_i-r_i$ is non-zero, that expression is negative.  
\end{enumerate}
\end{defn}
We leave it as an exercise to verify that this is an order.  \todo{check for ourselves that this actually is}
We now offer a lemma that allows us to inductively add $\mathbb{Q}$-points to a divisor and describe the new generators and relations.

@@

\todo{Peter:I'm still kind of unhappy with the notation $y_{e_i'}$ because it is treating $e_i'$ like a formal object rather than a number. What I mean is if $e_1'=e_2'$ that a reader might reasonably interpret $y_{e_1'}$ the same as $y_{e_2'}$}
\begin{enumerate}
	\item[(Ad-i)] First, 
	\begin{align*}
	\deg y_{e_i'} &= e_i'  &\text{ and } &&-\ord_{Q_i}^{\halfcan_X'}(y_{i, e_i'})
		= \frac{e_i'- 1}{2}.
	\end{align*}
		\item[(Ad-ii)] Second, every generator $z \neq y_{i, e_i'}$	of $R'$ satisfies
		\begin{align*}
			\frac{-\ord_{Q_i}
^{\halfcan_X'}(z)}{\deg z} < \subhalf {e_i'}.
		\end{align*}
	\item[(Ad-iii)] Third, we have
		\begin{align*}
			\deg \lfloor e_i L \rfloor \geq 2g - 1 + \max_{d \geq 0} \# S_{\sigma, J}(i, d)
		\end{align*}
		where
		\begin{align*}
			S_{\sigma, J}(i, d) := \{j \in J : j \neq i \text{ and } e_j'+2d
			\mid e_i - e_j'\}
		\end{align*}
\end{enumerate}
\end{defn}

>>>>

Following Voight and Zureick-Brown \cite{vzb:stacky}, we generalize Petri's theorem in many
directions. 
If $X$ is a curve, then we say $\sx$ is a {\bf stacky curve} with {\bf coarse space} $X$ and {\bf stacky points} $P_1,\ldots, P_r$ with {\bf stabilizer orders} $e_1,\ldots, e_r \in \BZ_{\geq 2}$ if 
$$\di \sx = \di X \oplus \left(\bigoplus_{i
= 1}^r \left\langle \frac{1}{e_i}P_i \right\rangle \right) \subset \BQ \otimes \di X.$$
As is done in \cite{vzb:stacky} one may view a stacky curve as a stack, but for the purposes of this paper, it suffices to view $\sx$ as a scheme whose associated divisor group allows a finite number of ``fractional points,'' i.e., stacky points. A {\bf log spin curve} is a triple $(\sx, \Delta, \halfcan)$ where $\Delta \in \di X$ is an effective divisor and $2\halfcan = K_X + \Delta + \sum_{i=1}^{r}\subhalf{e_i}P_i.$
The central object of study in this paper is the {\bf log spin
canonical ring} of $(\sx,\Delta,L)$ is
\begin{align*}
	R(\sx,\Delta,L) = \bigoplus_{d \geq 0} H^0(\sx,\sco(L)^{\otimes d})
\end{align*}

@@

\todo{Somewhere in the paper, we should probably give an example showing the generators and relations are sharply bounded at degree e, 2e. Robin: Aren't the base cases suitable examples? Aaron: Those don't show it for general e, do they? I don't think we really need an example, just a remark somewhere that one cannot really hope to do better than e, 2e somewhere.}

\section{Background}
\label{sec:background}
Look at Spencer-Kodaira \cite{kodaira:complex-manifolds}.
Include the following:
\begin{enumerate}
	\item Possibly include some of the following notation, it it isn't dealt with well in the background.

 A divisor $\Delta \in \di X$ is a {\bf log divisor} if
$\Delta = \sum_{i=1}^{n}P_i \in \di X,$ with $P_i$ distinct points of $X$. \todo{Do we want to assume they are distinct?} Let
$\sx$ denote a tame separably rooted curve over a
field $k$ so that $X$ is the coarse space of $\sx$, as defined in
\cite[Section 5.2]{vzb:stacky}, with $\di \sx = \di X \oplus \left(\bigoplus_{i
= 1}^r \langle \frac{1}{e_i}P_i \rangle \right),$ so that $e_i \in \BZ$. Let $K_
\sx = K_X + \sum_{i = 1}^{r}\frac{e_i - 1}{e_i}P_i.$ In this case, we
say $\sx$ has {\bf stabilizer order} $e_i$ at $P_i$. Call $\halfcan
\in \di \sx$ a {\bf log half-canonical divisor}, also known as a theta-characteristic, on the {\bf log stacky curve} $(\sx,\Delta)$ if
$2\halfcan \sim K_\sx + \Delta$. If $X$ has genus $g$ and $\deg \Delta =\delta,$ then we
say $\sx$ has \textbf{signature} $\sigma = (g; e_1, \ldots, e_r; \delta)$.

	\item Fix notation $\delta$ for $\deg \Delta$, $g$ for genus.
	\item While the remainder paper could be writen using stack-theoretic terms, as is done in \cite{vzb:stacky}, we shall adopt the simpler language of $\BQ$ divisors, thinking of $\sx$ as the underlying scheme $X$ with fractional points.
	\item Include explicit definitions for what we are assuming about a curve, i.e. smooth, projective, etc. Also, possibly discuss the relations to stacks, either here in or in the introduction. Make sure to make clear the assumption that we will work over an algebraically closed field through tensoring up.
	\item Notation for Coarse space
	\item Notation for Stacky points with associated stabilizer orders (I.e. as those points for which the divisor can have rational coefficients with denominator equal to the stabilizer order)
	\item Note that this result is independence of base field
	\item background on Gr\"{o}bner bases
	\item Notation $h^0 = \dim_k H^0$.
	\item Why we have these fractions $\frac{e_i - 1}{2e_i}$ and why $e_i$ must be odd.
	\item Include notation for $X$ as coarse space and $L_X$ the corresponding log half canonical divisor.
	\item Notation for Gr\"{o}bner bases and orderings (include grlex, grevlex, and block order, as defined in \cite[Definition 8.1.1]{vzb:stacky})
	\item Make sure to standardize the notation $\sx,\sx'$ for stacky curves, $L,L'$ for half canonical divisors, and $X,X'$ for (non-stacky) curves.
\end{enumerate}

@@

In Lemma \ref{lem:sat-1} we provide an inductive method to strengthen
O'Dorney's result \todo{cite theorem 8, Aaron: If you want to cite Theorem 8, just do \cite[Theorem 8]{dorney:canonical}} on $\BP^1_k$ to divisors $D = \sum_{i=1}^n \frac{\alpha_i
}{\beta_i} P_i$ for points $P_1, \ldots, P_n$ on a curve $X$ such that $H^
0(X, \lfloor D \rfloor) > 0$; our result implies that $R_D$ is generated in
degrees at most $e = \max(\beta_i)$ with relations generated in
degrees at most $2e$.  Furthermore, given additional constraints on $H^0(
X, \lfloor dD \rfloor)$ for various values of $d$, we can apply the
same Riemann-Roch arguments in arbitrary genus. 
To do this, we extend an inductive result by Voight and
Zureick-Brown on stacky canonical rings to arbitrary $\BQ$-
divisors. \todo{cite theorem 8.3.1}

@@

As suggested by the previous lemma, we will need an ordering on monomials, and the following definition will prove useful in the ensuing lemma.
\begin{defn}
We equip the graded ring $k[x_1, \ldots x_m]$, with weight $a_i$ and pole order $-\ord_P(x_i)$ attached to $x_i$, the graded $P$ reverse lexicographic order {\bf{g-P-revlex}} if for $f=\prod_{i=1}^m {x_i}^{q_i}\ne g=\prod_{i=1}^m {x_i}^{r_i}$, $f\prec g$ iff one of the following holds
\begin{enumerate}
\item $\deg(f) < \deg(g)$
\item $\deg(f) = \deg(g)$ and $-\ord(f) > -\ord(g)$
\item $\deg(f) = \deg(g)$, $-\ord(f) = -\ord(g)$, and for $i$ maximal such that $q_i-r_i$ is non-zero, that expression is negative.  
\end{enumerate}
\end{defn}
We leave it as an exercise to verify that this is an order.  \todo{check for ourselves that this actually is, Aaron: I suggest not to use the phrase ``leave as an exercise'' in a paper}
We now offer a lemma that allows us to inductively add $\mathbb{Q}$-points to a divisor and describe the new generators and relations.

@@

\todo{Peter:I'm still kind of unhappy with the notation $y_{e_i'}$ because it is treating $e_i'$ like a formal object rather than a number. What I mean is if $e_1'=e_2'$ that a reader might reasonably interpret $y_{e_1'}$ the same as $y_{e_2'}$. Aaron: By all means, please suggest better notation and we can discuss it. However, I think your particular concern is well answered because we include the further subscript $i$. That is, $y_{1,e_1} \neq y_{2,e_2}$ even if $e_1 = e_2=5$, because $y_{1,5}$ and $y_{2,5}$ are different variables. Another thing to note here is that $e_i' = e_i + 2,$ so that the two will never be equal. Does this assuage your concern? Also, Further, when I read David's paper, he used this notation, and I was never confused by this notation.}
\begin{enumerate}
	\item[(Ad-i)] First, 
	\begin{align*}
	\deg y_{e_i'} &= e_i'  &\text{ and } &&-\ord_{Q_i}^{\halfcan_X'}(y_{i, e_i'})
		= \frac{e_i'- 1}{2}.
	\end{align*}
		\item[(Ad-ii)] Second, every generator $z \neq y_{i, e_i'}$	of $R'$ satisfies
		\begin{align*}
			\frac{-\ord_{Q_i}
^{\halfcan_X'}(z)}{\deg z} < \subhalf {e_i'}.
		\end{align*}
	\item[(Ad-iii)] Third, we have
		\begin{align*}
			\deg \lfloor e_i L \rfloor \geq 2g - 1 + \max_{d \geq 0} \# S_{\sigma, J}(i, d)
		\end{align*}
		where
		\begin{align*}
			S_{\sigma, J}(i, d) := \{j \in J : j \neq i \text{ and } e_j'+2d
			\mid e_i - e_j'\}
		\end{align*}
\end{enumerate}
\end{defn}
